Addition and removal energies of circular quantum dots

### 2017-04-24 Manuscript/paper.tex
Update paper

@@

@@

@@

\begin{figure}
  \centering
  \includegraphics{fig-gs2}
  \caption{Ground state energies for a selection of quantum dot parameters, plotted against the number of shells.}
  \label{fig:gs}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{fig-rel-slopes2}
  \caption{The impact of the interaction on convergence of addition and removal energies using IM-SRG + QDPT3.  For clarity, the plot does not distinguish between addition and removal energies.  The horizontal axis shows quantum system parameters, where $N$ is the number of particles and $\omega$ is the oscillator frequency.  The vertical axis shows $|\rho_{15}|$ (``relative slope''), which estimates the rate of convergence at 15 total shells.  The lower the value of $|\rho_{15}|$, the faster the convergence.  The data points are categorized by the interactions.  The trends suggest that the singular short-range part of the interaction has a much stronger impact on the convergence than the long-range tail.}
  \label{fig:rel-slopes}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{fig-by-num-shells-6-2-ground-normal}
  \caption{Ground state energy for quantum dots with 6 electrons plotted against $K$, the number of shells in the basis.}
  \label{fig:by-num-shells-6-2-ground-0-normal}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{fig-by-freq-10-6-normal}
  \caption{The behavior of ground state, addition, and removal energies as a function of the oscillator frequency $\omega$, with 10 shells in the basis.  The energy is normalized with respect to the Hartree--Fock values to magnify the differences.  Lower frequency leads to stronger correlations and thereby a more difficult problem.  The fact that CCSD and MP2 are similar is likely coincidence as they cross at around 12 shells and then diverge again (not shown in this plot).}
  \label{fig:by-freq-10-6-normal}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{fig-fit-2-1p0-add}
  \caption{A five-point fit of the addition energies of the $(N, \omega) = (6, 1.0)$ system.  The grey shaded region contains the masked data points, which are ignored by the fitting procedure.  The left figure plots the central difference of the addition energies $\varepsilon^{(+)}$ with respect to the number of shells $K$.  On such a plot, the power law model should appear as a straight line.  The fits are optimized using the procedure described in Section \ref{subsec:extrapolation}.  Note that the lines do not evenly pass through the points in the left plot as the fitting weights are tuned for the energy on a linear scale, not the energy differences on a logarithmic scale.}
  \label{fig:by-fit-2-1p0-add}
\end{figure}

There is large amount of flexibility in the application of many-body methods.  The approach we use is shown in Figure \ref{fig:methods}.  Such an ordering maximizes the benefits of each method: Hartree--Fock acts as an initial, crude procedure to ``soften'' the Hamiltonian, followed by IM-SRG or CC to refine the ground state energy, and then finally QDPT or EOM to refine the addition and removal energies.

The arrows in the diagram indicate the flow of numerical data, which are usually matrix elements of some many-body operator.  For HF and IM-SRG, they are simply (evolved) Hamiltonian matrix elements.  For CC, they would be elements of the cluster operator $\hat{T}$.

There are several paths in this diagram through which one can traverse to obtain the output observables.  Among these possibilities, we shall primarily focus on the three combinations: (a) HF + IM-SRG(2) + QDPT3, (b) HF + IM-SRG(2) + EOM, and (c) HF + CCSD + EOM.

Moreover, one may also omit some steps of the process.  For example, one can omit HF, but continue with the remaining two steps.  While this is doable, from our experience Hartree--Fock significantly improves the results of the later post-HF methods at very low cost compared to the post-HF methods.  Therefore, in practice there is little reason to omit HF.  We will however investigate the effects of removing one or more of the post-HF methods.

The addition and removal energies are similar, but they require an additional parameter: the total orbital angular momentum $M_\ell$, defined as the sum of the $m_\ell$ of each particle.  This is due to the presence of multiple states with near-degenerate energies.  For this paper, we will consider exclusively the addition/removal energies with the lowest $|M_\ell|$ subject to the constraint that the particle added/removed lies within the next/last shell.  In particular, this means the $N + 1$ states of interest are those with $|M_\ell| = K_{\mathrm{F}} \bmod 2$ where $K_{\mathrm{F}}$ is the number of occupied shells, while the $N - 1$ states of interest are those with $|M_\ell| = 1 - (K_{\mathrm{F}} \bmod 2)$.

All calculations in this paper are performed with a relative numerical precision of about $10^{-5}$ or lower (sometimes \emph{much} lower, as is for CC).  This does not necessarily mean the results are as precise as $10^{-5}$, since numerical errors tend to accumulate at every step of the calculation, thus the precision of the final results is more like $10^{-4}$.

In Figure \ref{fig:gs} we display a selection of ground state energies calculated using the three approaches described in Section \ref{subsec:methodology}.  We also include results from M\o ller--Plesset perturbation theory to second order (MP2), DMC \cite{hoegberget2013thesis}, and FCI \cite{olsen2013thesis} for comparison.

Both perturbation theory and IM-SRG add significant corrections to the Hartree--Fock results.  Compared with MP2, IM-SRG generally converges faster, although does not always converge to the DMC result.  Generally speaking, IM-SRG is closer to the DMC result as the number of particles increases.

There are a few cases where the IM-SRG over-corrects the result, leading to a ground state energy that is lower than the exact result.  This is not unexpected given that, unlike Hartree--Fock, IM-SRG is non-variational in the presence of operator truncations, which cause small unitarity violations.  Overcorrection tends to occur when the number of particles is few, or the frequency is very low (high correlation).  However, MP2 does not do much better in the high frequency regimes as it often converges slower than IM-SRG.

The operator truncations in the evaluation of the commutator is one of the two sources of discrepancy between IM-SRG and exact results.  The other source is that of basis truncation, which limits the size of the Hilbert space that can be explored by the method.  Unlike operator truncation, which is difficult to overcome due to its exponentially increasing computational cost, the cost of increasing the basis is ``only'' polynomial with respect to the number of basis elements: roughly $O(N_{\mathrm{b}}^6)$ where $N_{\mathrm{b}}$ is the total number of states in the single-particle basis.  In practice, the scaling cost is usually lower as the conserved symmetries allow the Hamiltonian to be partitioned into blocks.

For our addition and removal energy calculations, the results are summarized in Figure \ref{fig:add} and Figure \ref{fig:rm} respectively.  The figures show the the addition/removal energies for using the approaches mentioned in Section \ref{subsec:methodology}.  Where available, results from diffusion Monte Carlo (DMC) \cite{PhysRevB.84.115302} are shown as a dashed line.

From these results, we see that the QDPT corrections add a substantial contribution to the results of both Hartree--Fock with and without IM-SRG.  [[quantify ``substantial'' numerically]]

Interestingly, the addition energy of Hartree--Fock with second-order QDPT appears to surpass the accuracy of that of IM-SRG with second-order QDPT.  This is likely coincidental, as the third-order QDPT dramatically worsens the result.

Both Hartree--Fock and IM-SRG results worsen as the frequency decreases, which is not unexpected since the correlations become much more dominant.  The Hartree--Fock appears to perform much worse, however, as the results converge much more slowly as the number of shells increases.

We also note that IM-SRG with 2 particles and a frequency 0.1 generally fail to give any result as the ODE solver eventually leads to divergence [[Why?]]  This seems to be a quirk isolated to this particular system and does not usually occur in other systems.

For addition and removal energies, we found that IM-SRG improves their rate of convergence both as the order of perturbation theory increases and as the number of shells increases.

[[---]]

When the tail part of the convergence plots are magnified about a factor of ten or so [[double-check this number]], we noticed that the addition and removal energies are still drifting somewhat [[quantify how much]], hinting that the convergence rate is probably not exponential.

To investigate the rate of convergence, we tested our methods against a modified form of the Coulomb interaction, parametrized by two lengths $\sigma_{\mathrm{A}}$ and $\sigma_{\mathrm{B}}$ that characterize the range of the interaction:
\begin{align}
  V_{\sigma_{\mathrm{A}}, \sigma_{\mathrm{B}}}(r) \equiv \frac{(1 + c)^{1 - 1/c}}{c} \left(1 - \E^{-r^2 / (2 \sigma_{\mathrm{A}}^2)}\right) \E^{-r^2 / (2 \sigma_{\mathrm{B}}^2)} \frac{1}{r}
\end{align}
where $c \equiv \sqrt{\sigma_{\mathrm{B}} / \sigma_{\mathrm{A}}}$.  The coefficient is defined to ensure the peak of the envelope remains at unity.  By setting $\sigma_{\mathrm{A}} = 0$ and $\sigma_{\mathrm{B}} = \infty$ one recovers the original Coulomb interaction.  By increasing $\sigma_{\mathrm{A}}$ one can reduce the short-range part of the interaction, and analogously by increasing $\sigma_{\mathrm{B}}$ one can reduce the long-range part of the interaction.  For our numerical experiments we consider the following four combinations of $(\sigma_{\mathrm{A}}, \sigma_{\mathrm{B}})$: $(0, \infty)$, $(\frac{1}{2}, \infty)$, $(0, 4)$, $(\frac{1}{2}, 4)$.

To analyze the convergence as function of the number of shells $K$, we use a simple but approximate metric, the \textit{relative slope} at $K = 15$:
\begin{align*}
  \rho_{15}(\varepsilon) \equiv \frac{\varepsilon_{K = 15} - \varepsilon_{K = 14}}{\varepsilon_{K = 15}}
\end{align*}
For the systems we have selected, at $K = 15$ the addition and removal energies have essentially reached the asymptotic behavior (by visual inspection), thus $\rho_{15}$ gives a crude measure of the rate of convergence as a relative quantity independent of the energy unit, allowing it to be meaningfully compared between systems with different interactions and different particle numbers.  The results are plotted in Figure \ref{fig:rel-slopes}.

Analyzing the figure, we observe that the reducing short-range part of the interaction improves the rate of convergence greatly -- the relative slope has saturated and reached the precision of the ODE solver -- whereas eliminating the long-range part of the interaction had very little effect.  This demonstrates that the main cause of the slow convergence lies in the highly repulsive, short-ranged part of the interaction, which leads to the presence of nondifferentiable cusps (the so-called \textit{Coulomb cusps}) in the exact wave functions that are difficult to reproduce exactly using linear combinations of the smooth harmonic oscillator wave functions.

The figure also tells us that the convergence is negatively impacted at lower frequencies and, to a lesser extent, by the number of particles.  Both are expected: lower frequencies increase the correlation in the system, while higher number of particles naturally require more shells to converge.

We also note that the behavior of addition energies versus removal energies tend to be similar, and which one converges faster fluctuates from system to system.

As suggested by Kvaal \cite{PhysRevB.80.045321,Kvaal2007}, the asymptotic convergence of the wave function can be approximately described by a power law model: [[assuming I understood his paper correctly]]
\begin{align*}
  \Delta E \propto K^{-\beta}
\end{align*}
where $\Delta E$ is the error in the energy, $K$ is the number of shells in the single-particle basis, and $\beta$ is some positive real exponent.  Essentially, the \textit{smoothness} of the exact wave function determines the rate of the convergence: the more times the exact wave function can be differentiated, the higher the exponent $\beta$.

It is difficult to ascertain what $\beta$ is \textit{a priori}, but based on Kvaal's results $\beta$ is expected to be below $5$ for quantum dot systems.  To determine this from our numerical calculations, we perform a curve using the following power-law model:
\begin{align} \label{eq:power_law_model}
  E = \alpha K^{-\beta} + \gamma
\end{align}
As with most nonlinear curve fits, it is often fairly sensitive to the initial parameters.  To extract the initial parameters, we first do a linear fit on $\ln (\partial E / \partial K)$ against $\ln (K)$:
\begin{align*}
  \ln \left(\frac{\partial E}{\partial K}\right) = \ln(\alpha) - \beta \ln(K)
\end{align*}
The derivative is approximated using finite differences of adjacent data points:
\begin{align*}
  \frac{\partial E}{\partial K} \approx E\left(K + \frac{1}{2}\right) - E\left(K - \frac{1}{2}\right)
\end{align*}
The process of calculating a derivative amplifies the noise, but linear fits do have the strong advantage of being very reliable to fit -- there is little concern about being stuck in a local minimum -- as well as making the results easy to judge by visually.  Of course, linear fits do not weigh the individual contributions equally, thus a more accurate nonlinear curve fit is required afterward.

We use the results as inputs for a power-law fit of $E$ against $K$, which is not as sensitive to noise.  We do require a guess for the infinite-basis energy $\gamma$, which we estimate by fitting \eqref{eq:power_law_model} while the parameters $\alpha$ and $\beta$ are fixed to the initial guesses.  Finally, we do a fit with all three parameters free to vary.

There is still one additional tuning knob for this model that is not explicitly shown in \eqref{eq:power_law_model}: the range of data points taken into consideration.  Since the model describes the \emph{asymptotic} behavior, we do not expect the fit to produce good results when the energy is still far from convergence.  To remedy this, we restrict ourselves to fitting only the last few data points within some chosen range.  If range is too large, then the non-asymptotic behavior would perturb the result too much, whereas if the range is too small, the noise will obscure the result.  This part is tuned by hand based on observations of the log-log plot: we expect the asymptotic region to lie more or less on a straight line.

[[Idea for further exploration: see how much the extrapolation improves the result, and see what happens as we increase the number of shells included in the fit on the right/upper bound.]]

Unlike all of the other methods, Hartree--Fock energies are somewhat unusual in that they generally do not fit to well to the power-law model.  In fact, the results and plots seem to indicate an \emph{exponential} convergence with respect to the number of shells.  Based on the results with artificial softened interactions as well as past experiences, we believe this is an artifact of the insensitivity of Hartree--Fock theory to the Coulomb cusp.

Nonetheless, despite the poor fits that often arise, the extrapolated energies are often quite good for HF.  This is likely due to its rapid convergence, which leaves very little degree of freedom even for a badly chosen model.  Moreover, we find that the fit uncertainties of energy are also roughly within the right order of magnitude, surprisingly.

>>>>

There is great amount of flexibility in the application of many-body methods.  The approaches we use is depicted in Figure \ref{fig:methods}.  Such an ordering maximizes the benefits of each method: Hartree--Fock acts as an initial, crude procedure to ``soften'' the Hamiltonian, followed by IM-SRG or CC to refine the ground state energy, and then finally QDPT or EOM to refine the addition and removal energies.

The general process begins with the input matrix elements \eqref{eq:interactionintegral}, computed exactly using the \texttt{OpenFCI} software library \cite{2008arXiv0810.2644K}.  Internally, \texttt{OpenFCI} calculates the interaction integrals in the center-of-mass frame using Gauss--Hermite quadrature and transforms them back into the laboratory frame to produce matrix elements suitable for use in our many-body methods.

Afterward, there are several paths through which one can traverse Figure \ref{fig:methods} to obtain output observables.  We shall primarily focus on the three combinations: (a) HF + IM-SRG(2) + QDPT3, (b) HF + IM-SRG(2) + EOM2, and (c) HF + CCSD + EOM2.

It is possible to omit some steps of the process.  For example, one can omit HF, but continue with the remaining two steps.  While this is doable, from our experience Hartree--Fock significantly improves the results of the later post-HF methods at very low cost compared to the post-HF methods.  Therefore, in practice there is little reason to omit HF.  We will however investigate the effects of removing one or more of the post-HF methods.

The addition and removal energies are similar, but they require an additional parameter: the total orbital angular momentum $M_\ell$, defined as the sum of the $m_\ell$ of each particle.  This is due to the presence of multiple states with near-degenerate energies.  For this paper, we will consider exclusively the addition/removal energies with the lowest $|M_\ell|$ subject to the constraint that the particle added/removed lies within the next/last shell.  This means the $N + 1$ states of interest are those with $|M_\ell| = K_{\mathrm{F}} \bmod 2$ where $K_{\mathrm{F}}$ is the number of occupied shells, while the $N - 1$ states of interest are those with $|M_\ell| = 1 - (K_{\mathrm{F}} \bmod 2)$.

Numerical calculations in this paper are performed with a relative precision of about $10^{-5}$ or lower.  This does not necessarily mean the results are as precise as $10^{-5}$, since numerical errors tend to accumulate over the multiple steps of the calculation, thus the precision of the final results is expected to be roughly $10^{-4}$.

@@

\begin{figure}
  \centering
  \includegraphics{fig-gs2}
  \caption{Plots of ground state energy of quantum dots with $N$ particles and an oscillatory frequency of $\omega$ against the number of shells $K$.  Since DMC does not utilize a finite basis, the horizontal axis is irrelevant and DMC results are plotted as horizontal lines.}
  \label{fig:gs}
\end{figure}

In Figure \ref{fig:gs} we display a selection of ground state energies calculated using HF + CC and HF + CCSD as described in Section \ref{subsec:methodology}.  We also include results from M\o ller--Plesset perturbation theory to second order (MP2), DMC \cite{hoegberget2013thesis}, and FCI \cite{olsen2013thesis} for comparison where available.

We do not include results from ``HF only'' to avoid overshadowing the comparatively smaller differences between the non-HF results in the plots.  Some HF results can be found Figure \ref{fig:by-freq-10-6-normal} instead.  Generally, HF ground state energies differ from the non-HF ones by a few to several percent, whereas non-HF energies tend to differ from each other by less than a percent.

With respect to the number of shells, both IM-SRG(2) and CCSD appear to converge slightly faster than second order perturbation theory (MP2), likely due to the presence of higher order corrections in IM-SRG(2) and CCSD.

There are a few cases where the IM-SRG over-corrects the result, leading to an energy lower than the quasi-exact DMC results.  This is not unexpected given that, unlike Hartree--Fock, IM-SRG is non-variational in the presence of operator truncations, which cause small unitarity violations.  This over-correction tends to occur when the frequency is low (high correlation), or somewhat counterintuitvely, when \emph{few} particles are involved.

@@

The results of our addition and removal energy calculations are summarized in Figure \ref{fig:add} and Figure \ref{fig:rm} respectively.  The figures show the the addition/removal energies for using the approaches mentioned in Section \ref{subsec:methodology}.  Where available, results from diffusion Monte Carlo (DMC) \cite{PhysRevB.84.115302} are shown as a dashed line.

As before, we do not include results from ``HF only'' in these plots as they are significantly further from the rest.  Analogously, we also exclude results from pure IM-SRG (i.e.\ without QDPT nor EOM) or pure CCSD, as QDPT or EOM both add significant contributions to addition and removal energies.  Some HF only and pure IM-SRG results can be seen in Figure \ref{fig:by-freq-10-6-normal}.

There is strong agreement between IM-SRG(2) + QDPT3 and IM-SRG(2) + EOM2 in many cases, and slightly weaker agreement between the IM-SRG and CCSD families.  This suggests that the EOM2 corrections are largely accounted for by the inexpensive QDPT3 method.  However, in some cases, most notably with few particles and high correlations (low frequency), the IM-SRG(2) + QDPT3 result differs significantly from both IM-SRG(2) + EOM2 and CCSD + EOM2.

\begin{figure}
  \centering
  \includegraphics{fig-by-freq-10-6-normal}
  \caption{The behavior of ground state, addition, and removal energies as a function of the oscillator frequency $\omega$, with $K = 10$ shells in the basis.  The energy is normalized with respect to the Hartree--Fock values to magnify the differences.  Lower frequency leads to stronger correlations and thereby a more difficult problem.  The fact that CCSD and MP2 are similar is likely coincidence as they cross at around 12 shells and then diverge again (not shown in this plot).}
  \label{fig:by-freq-10-6-normal}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{fig-rel-slopes2}
  \caption{The impact of the interaction on convergence of addition and removal energies using IM-SRG(2) + QDPT3.  For clarity, the plot does not distinguish between addition and removal energies.  The horizontal axis shows the system parameters, where $N$ is the number of particles and $\omega$ is the oscillator frequency.  The vertical axis shows $|\rho_{15}|$ (``relative slope''), which estimates the rate of convergence at 15 total shells.  The lower the value of $|\rho_{15}|$, the faster the convergence.  The data points are categorized by the interactions.  The trends suggest that the singular short-range part of the interaction has a much stronger impact on the convergence than the long-range tail.}
  \label{fig:rel-slopes}
\end{figure}

To analyze the rate of convergence more quantitatively, we shall define $\rho_K$ as the relative backward difference of the energy (\textit{relative slope}):
\begin{align*}
  \rho_K = \frac{\varepsilon_{K} - \varepsilon_{K - 1}}{\varepsilon_K}
\end{align*}
The denominator allows the quantity to be meaningfully compared between different systems.  We expect this quantity to become increasingly small as the calculations converge.

In Figure \ref{fig:rel-slopes}, we plot the $\rho_{15}$ for IM-SRG(2) + QDPT3.  The many-body methods were tested against a \emph{modified} Coulomb-like interaction, parametrized by two lengths $\sigma_{\mathrm{A}}$ and $\sigma_{\mathrm{B}}$ that characterize the range of the interaction:
\begin{align}
  V_{\sigma_{\mathrm{A}}, \sigma_{\mathrm{B}}}(r) \equiv \frac{(1 + c)^{1 - 1/c}}{c} \left(1 - \E^{-r^2 / (2 \sigma_{\mathrm{A}}^2)}\right) \E^{-r^2 / (2 \sigma_{\mathrm{B}}^2)} \frac{1}{r}
\end{align}
where $c \equiv \sqrt{\sigma_{\mathrm{B}} / \sigma_{\mathrm{A}}}$.  The coefficient is chosen to ensure the peak of the envelope remains at unity.  With $(\sigma_{\mathrm{A}}, \sigma_{\mathrm{B}}) = (0, \infty)$ one recovers the original Coulomb interaction.  By increasing $\sigma_{\mathrm{A}}$ one can truncate the short-range part of the interaction, and analogously by increasing $\sigma_{\mathrm{B}}$ one can truncate the long-range part of the interaction.  For our numerical experiments we considered the following four combinations of $(\sigma_{\mathrm{A}}, \sigma_{\mathrm{B}})$: $(0, \infty)$, $(\frac{1}{2}, \infty)$, $(0, 4)$, $(\frac{1}{2}, 4)$.

Reducing short-range part of the interaction appears improves the rate of convergence substantially.  Many of the cases have reached the precision of the ODE solver ($10^{-5}$ to $10^{-6}$).  In contrast, eliminating the long-range part of the interaction had very little effect.  This suggests that the main cause of the slow convergence lies in the highly repulsive, short-ranged part of the interaction, which leads to the presence of nondifferentiable cusps (the so-called \textit{Coulomb cusps}) in the exact wave functions that are difficult to reproduce exactly using linear combinations of the smooth harmonic oscillator wave functions.

The convergence is negatively impacted at lower frequencies and, to a lesser extent, by the increased number of particles.  Both are expected: lower frequencies increase the correlation in the system, while higher number of particles naturally require more shells to converge.

In general, there does not appear to be any difference between the convergence behavior of addition energies as compared to that of removal energies.

@@

\begin{figure}
  \centering
  \includegraphics{fig-fit-2-1p0-add}
  \caption{A five-point fit of the addition energies of the $(N, \omega) = (6, 1.0)$ system with $K_{\text{stop}} = 15$.  The grey shaded region contains the masked data points, which are ignored by the fitting procedure.  The left figure plots the central difference of the addition energies $\varepsilon^{(+)}$ with respect to the number of shells $K$.  On such a plot, the power law model should appear as a straight line.  The fits are optimized using the procedure described in Section \ref{subsec:extrapolation}.  Note that the lines do not evenly pass through the points in the left plot as the fitting weights are tuned for the energy on a linear scale, not the energy differences on a logarithmic scale.}
  \label{fig:by-fit-2-1p0-add}
\end{figure}

As derived by Kvaal \cite{PhysRevB.80.045321,Kvaal2007}, the asymptotic convergence of the quantum dot observables in a finite harmonic oscillator basis can be approximately described by a power law model:
\begin{align*}
  \Delta E \propto K^{-\beta}
\end{align*}
where $\Delta E$ is the difference between the finite-basis result and the infinite-basis result, $K$ is the number of shells in the single-particle basis, and $\beta$ is some positive real exponent.  The \textit{smoothness} of the exact wave function determines the rate of the convergence: the more times the exact wave function can be differentiated, the higher the exponent $\beta$.

It is difficult to ascertain what $\beta$ is \textit{a priori}, thus we will empirically compute $\beta$ by fitting the following model through our data:
\begin{align} \label{eq:power_law_model}
  E = \alpha K^{-\beta} + \gamma
\end{align}
As a nonlinear curve fit, it can be quite sensitive to the initial parameters.  Therefore, good guesses of the parameters are necessary to obtain a sensible result.  For this, we first fit a linear model of $\log |\partial E / \partial K|$ against $\log K$:
\begin{align*}
  \log \left|\frac{\partial E}{\partial K}\right| = - (\beta + 1) \log K + \log|\alpha \beta|
\end{align*}
This is useful because linear fits are very robust and will often converge even if the initial parameters are complete nonsense.  It also provides a means to visually judge quality of the fit.  The derivative is approximated using the central difference:
\begin{align*}
  \frac{\partial E}{\partial K} \approx E\left(K + \frac{1}{2}\right) - E\left(K - \frac{1}{2}\right)
\end{align*}
The process of calculating a numericaly derivative can amplify the noise in the data and distorts the weights of the data points.  Moreover, it does not provide a means to compute $\gamma$, the extrapolated energy.  Thus a second accurate nonlinear curve fit is necessary.

The parameters $\alpha$ and $\beta$ are extracted from the linear fit and used as inputs for a power-law fit of $E$ against $K$.  It necessary to estimate the infinite-basis energy $\gamma$ as well, which is done by fitting \eqref{eq:power_law_model} while the parameters $\alpha$ and $\beta$ are fixed to the initial guesses.  The fixing ensures that the fit is still linear in nature and thus highly likely to converge.  Afterward, we do a final fit with all three parameters free to vary.  All fits are done using the traditional Levenberg--Marquardt (LM) optimization algorithm as implemented in \texttt{Minpack}, with equal weighting of all data points.

There is still one additional tuning knob for this model that is not explicitly part of \eqref{eq:power_law_model}: the range of data points taken into consideration (\textit{fit range}).  Since the model describes the \emph{asymptotic} behavior, we do not expect the fit to produce good results when the energy is still very far from convergence.  To account for this, we only fit the last few data points within some chosen range.  If range is too large, then the non-asymptotic behavior would perturb the result too much, whereas if the range is too small, there would be more noise and less confidence in whether the trend is legitimate rather than accidental.  Empirically, we chose to fit the last 5 points of our available data.  The results are shown in Tables \ref{tab:ground-extrapolated}, \ref{tab:add-extrapolated}, and \ref{tab:rm-extrapolated}.  A specific example of the fit is shown in Figure \ref{fig:by-fit-2-1p0-add}

The LM fitting procedure also computes uncertainties for the parameters from an approximate Hessian of the model function.  It is therefore tempting to use the uncertainty of the fit to quantify the uncertainty of the extrapolated energy.  We certainly would not expect this to account for the error due to the operator truncation, but how accurately does it quantify the discrepancy of our extrapolated result from the true infinite-basis energy?

We investigated this idea by performing a fit over \emph{all possible} 5-point fit ranges $[K_{\text{stop}} - 4, K_{\text{stop}}]$.  By comparing the extrapolated results at varying values of $K_{\text{stop}}$ with the extrapolated result at the highest possible $K_{\text{stop}}$ and treating the latter as the ``true'' infinite-basis result, we can statistically assess whether the fit uncertainties are a good measure of the discrepancy from the true infinite-basis result.  Our results show a somewhat bimodal distribution: when the relative fit uncertainty is higher than $10^{-3.5}$, the fit uncertainty quantifies the discrepancy well; otherwise, the fit uncertainty underestimates the discrepancy by a factor of 10 or less.

Unlike the other methods, Hartree--Fock energies are somewhat unusual in that they generally do not conform to the power-law model.  In fact, the results and plots seem to indicate an \emph{exponential} convergence with respect to the number of shells.  It is likely that HF is insensitive to the Coulomb cusp.

Nonetheless, despite the poor fits that often arise, the extrapolated energies are often quite good for HF.  This is likely due to its rapid convergence, which leaves very little degree of freedom even for a poorly chosen model.  Moreover, we found that the fit uncertainties of energy are fairly good measures of the true discrepancy.

Not all fits yield a positive value of $\beta$ for addition and removal energies, which suggests that the data points do not converge, or require a very high number of shells to converge.  These cases are marked ``n.f.'' in the tables.  This affects exclusively IM-SRG(2) + QDPT3 for systems with few particles and low frequencies, indicating that perturbation theory is inadequate for such systems.
